\vspace{-2cm}A generative model for urban firm networks

### 2022-05-23 S2_Text.tex
Update on Overleaf.

\section*{S2 Text: Statistical models, generative model setup and sensitivity analysis}

@@

>>>>

\section*{S2 Text: Statistical models, generative model setup, sensitivity analysis, and model calibration}

@@

\subsection*{Model calibration}

When all parameters except $w_0$ are fixed, an external constraint can be enforced, hence fixing its value. A simple example is to set the total volume of simulated links to that it equals the total volume of observed links. Since by construction we have at final time $\sum_{i,j} w_{ij} = w_0 \cdot t_f$, this constraint yields $w_0 = \frac{\sum_{ij} w_{ij}^{(obs)}}{t_f}$ with $w_{ij}^{(obs)}$ observed weights.

For the model calibration, we choose a constraint minimising one of the two objectives. Rescaling the final simulated flows by some constant $k_0$ while having $w_0 = 1$ is equivalent to set $w_0 = k_0$. We can thus determine the optimal value of $k_0$ that will minimise

\[
\varepsilon_L (k_0) = \frac{1}{N^2} \sum_{i,j} \left(\log w_{ij} - \log \left(k_0 \cdot \hat{w}_{ij} \right) \right)^2
\]

The derivative along $k_0$ must thus be zero, what yields

\[
0 =  \frac{1}{N^2} \sum_{i,j} \log w_{ij} - \frac{1}{N^2} \sum_{i,j} \log \hat{w}_{ij} - \frac{1}{N^2} \sum_{i,j} \log k_0
\]

and therefore

\[
k_0 = \exp \left[\frac{1}{N^2} \sum_{i,j} \log w_{ij} - \frac{1}{N^2} \sum_{i,j} \log \hat{w}_{ij} \right]
\]

In practice, we rescale the final simulated flows by this value in the calibration procedure. Note that only one constraint can be satisfied, and that this rescaling which improves the $\varepsilon_L$ objective probably deteriorates the $\varepsilon_M$ objective. We choose to rescale this dimension as it is directly comparable with the statistical models.

### 2022-05-23 main.tex
Update on Overleaf.

\section*{Abstract}
The emergence of interconnected urban networks is a crucial feature of globalisation processes. Understanding the drivers behind the growth of such networks -- in particular urban firm networks --, is essential for the economic resilience of urban systems. We introduce in this paper a generative model for firm networks at the urban area level, which considers the following features: the economic size of the areas representing the source and target nodes, the industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distances between urban areas. An empirical network analysis on European urban firm ownership data confirms the relevance of each of these factors on new link creation. We then simulate network growth on synthetic systems of cities, unveiling stylized facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range. We calibrate the model on the European network, outperforming statistical models, and showing a strong role of path-dependency. The model is applied to stylised scenarios, opening up to the formulation of public policies recommendations to deal with exogenous shocks such as economic crises or potential lock-downs of countries.

@@

Networks constitute the new social morphology of our societies, since the intensification of globalisation of the economy in the late 20th century \cite{castells2000networksociety}. A world-city system has since then emerged \cite{taylor2001specification}, characterized by highly interconnected urban centres at the global level \cite{sassen1991}. These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity inducing spatial and non-spatial patterns \cite{martinus2018global}. Interactions of economic nature between transnational firms \cite{derudder2018central}, are together with international trade patterns among the most representative of the current economic trends \cite{taylor2001specification}. Understanding the drivers of growth and geographical properties of such an urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}. In addition, strengthening city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}. This can be achieved through strategic development of certain industrial sectors for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}. Exploring city/firm interactions \cite{storme2019introducing} and the position of cities within these networks \cite{gluckler2016relational} can also allow us to infer the consequences of unforeseen exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.

@@

At each time step, one directed link is randomly drawn. We define a generalized Cobb-Douglas utility function \cite{vilcu2011geometric} as

@@

The initial network is the same than in the synthetic setup, and the real links are taken as targets for the generated network. We constrain the total volume of flows by setting $w_0 = \sum_{ij} w_{ij}^{(obs)} / t_{f}$ when the number of time steps $t_f$ is given as a parameter.

@@

We base our study on empirical data from the \emph{AMADEUS} dataset produced by the Bureau Van Dijk (BVD) on transnational firms in Europe. \emph{AMADEUS} provides a precise location, turnover and activity codes for owners and owned companies in all sectors in Europe. The Bureau Van Dick gathers, homogenizes and completes the information on these firms from national sources, publications of consultancies and audit firms such as KPMG, Deloitte, PricewaterhouseCoopers (PwC), Ernst \& Young (EY), Mazars, and those of other financial organisations such as Moody's, Fitch, Standard \& Poor's, Thomson Reuters or Duns and Bradstreet. A representativity study shows that \emph{ORBIS} --the Bureau Van Dijk's dataset for firms at the worldwide scale--, captures around 10 percent of existing foreign direct investments in 2011 in Poland \cite{Zdanowska2018}. \cite{bajgar2020coverage} find that the coverage of the \emph{ORBIS} database on OECD countries is a representative sample, but with a better performance on larger firms. As we aggregate company turnovers on urban areas in our analysis, we thus do not expect much representativity bias in the inter-urban network.

@@

A global sensitivity analysis first unveils an aggregated influence of parameters on model indicators. We apply Saltelli's Global Sensitivity method introduced by \cite{saltelli2008global}. This produces values for each parameter and each output of a first order sensitivity index, defined as $Var\left[\mathbb{E}_{\mathbf{X}_{\sim i}}\left(Y_j | X_i\right)\right] / Var\left[Y_j\right]$, for parameter $X_i$ and output $Y_j$, $\mathbf{X}_{\sim i}$ being all other parameters. This captures the total influence of $X_i$ all other things being equal. The total order index is given by $\mathbb{E}_{\mathbf{X}_{\sim i}} \left[Var(Y_j | \mathbf{X}_{\sim i})\right] / Var\left[Y_j\right]$ and captures non-homogeneity in behaviour and interaction between parameters. Indices were estimated with a Sobol sequence sampling of 20,000 parameter points (corresponding to 160,000 model runs) \cite{saltelli2010variance}. We give in S2 Text the full estimated results for sensitivity indices. To summarize, we find that indicators related to modularity (internationalisation and optimal modularity) are mostly influenced by geographical parameters. On the contrary, indicators of network structure are driven by origin and destination sizes. The influence of sectors and historical links has a secondary but not negligible role. Altogether, the analysis shows strong interaction effects between parameters, since for several indicators the total order index is one order of magnitude larger than the first order index. This means that the generative model captures some aspects of the complex behaviour of the system.

@@

We run a specific experiment to study the impact of the urban hierarchy on firms dynamics. Rank-size law in cities is a central stylized fact within urban theories, despite the discussions regarding the estimation of the power-law exponent, its actual values and its variation depending on the definition of cities considered \cite{cottineau2017metazipf,corral2020truncated}. I this particular context, understanding how this parameter can impact urban dynamics within artificial systems of cities -- in which it can be modified to extreme values non observable in real systems--, should help disentangling the variability due to model parameters from the variability due to the specific configuration of the urban system \cite{raimbault2019space}. We change the initial hierarchical exponent $\alpha$ from 0.1 to 2.0 with a step of 0.1, and also vary the interaction range $d_0$. The behaviour of the internationalisation and metropolisation indicators are shown in Fig.~\ref{fig:fig7}. As expected, the hierarchy of the urban system has an important impact on model outcomes. We find that more hierarchical urban systems tend to produce more international networks (recall that the internationalisation indicator corresponds to the modularity, the latter is minimized when the networks span more across different countries). This is consistent with the existence of globalised metropolis \cite{sassen1991global}. This effect is mitigated by reducing the interaction ranges. Regarding metropolisation, we observe the existence of maxima as a function of the initial hierarchy across different interaction ranges $d_0$. This means that a too high hierarchy is detrimental to the global metropolisation when considering all cities, what may be due to the fact that in such a case very large cities are capturing all flows, leaving nothing to intermediate and medium-sized cities. Interestingly, the optimal value $\alpha \simeq 1.1$ is not far from the observed values (which usually range between 0.9 and 1.1 but can take larger values such as 1.5 when extending the urban system \cite{Raimbault_2020}), implying that this optimality property is endogenous to the emergence of urban systems and linked to their intrinsic hierarchy.

%%%%%%%%%%%%%
\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{figures/Fig7.png}
    \end{center}
    \caption{\textbf{Impact of the urban hierarchy.} (Left) Internationalisation as a function of the initial synthetic hierarchical exponent $\alpha$, for different values of the interaction range $d_0$ (color); (Right) Metropolisation as a function of $\alpha$, for different values of the interaction range $d_0$ (color).\label{fig:fig7}}
\end{figure}
%%%%%%%%%%%%%

@@

% TODO precise calib setting + details on rescaling
We then apply the model to a real system of cities by calibrating it on the European ownership network. The model setup is done following the real setup described above. The number of time steps $t_f$ is left as an additional parameter, which in a sense determines the granularity of the cumulative network generation process.

@@

%%%%%%%%%%%%%%
\begin{figure}
	\begin{center}
    \includegraphics[width=0.75\linewidth]{figures/Fig8.png}
    \end{center}
	\caption{\textbf{Model calibration results.} Pareto front obtained for the bi-objective model calibration for $\varepsilon_L$ and $\varepsilon_M$ objectives. $\varepsilon_L$ can be compared to the MSE of statistical models fitted above, and the generative model outperform these for several parametrisations (lowest MSE 5.33 for statistical models). Point size gives the number of stochastic samples and color the value of $\gamma_O$.\label{fig:fig8}}
\end{figure}
%%%%%%%%%%%%%%

@@

When considering optimal points with $\varepsilon_L < 5.0$, the adjusted parameters are: long range gravity interactions and socio-economic interactions given by $d_0 = 5972 \pm 2148$ and $c_0 = 113 \pm 33$ (the maximal $c_{ij}$ is 11.8); a stronger influence of the destination than of the origin as $\gamma_O = 1.46 \pm 0.10$ and $\gamma_D = 2.0 \pm 0.05$; a comparable influence of sectors as $\gamma_S = 1.76 \pm 0.21$; a fine time granularity as $t_f = 4951 \pm 124$; and a very strong influence of previous links as $\gamma_W = 6.43 \pm 4.11$. This last point furthermore shows in particular the importance of including path-dependency in this complex generative model. The interaction of this process with others implies an inversion of the relative importance of origin and destination (in opposition to the statistical models). The fact that calibrated interaction distances are large is consistent with long range economic interactions within Europe. Besides these stylized facts, these calibration results altogether demonstrate the feasibility of the application of the model to real data. As stated previously, the model is not predictive and the relatively high error values would not make it satisfactory in that manner, but are reasonable for its explanatory and application functions.

@@

%%%%%%%%%%%%%%
\begin{figure}
	\begin{center}
    \includegraphics[width=\linewidth]{figures/Fig9.png}
    \end{center}
	\caption{\textbf{Application of the calibrated model to an economic shock scenario.} All plots show the trajectory of indicators (internationalisation for the top row and average community size for the bottom row) in time, for different values of interaction decay rescaling $\kappa_G$ (color) and socio-economic distance rescaling $\kappa_C$ (panels). The vertical dotted red line shows the moment the economic shock is introduced and $d_0$ (respectively $c_0$) is set to $\kappa_G \cdot d_0$ (resp. $\kappa_C \cdot c_0$).\label{fig:fig9}}
\end{figure}
%%%%%%%%%%%%%%

@@

The model calibration results provide model parametrisations reasonably approaching the observed network. This validates its potential application to stylised policy issues with practical applications in the future. For example the effect of exogenous shocks in the socio-economic structure can be tested as we did in the stylized application example above. Such shocks can for example be the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to further pandemic outbreaks on the single market. In this sense the UK economic sectors mostly dependent on foreign capital and generating the highest turnovers in 2018, that would be the most affected and worth examining in detail are: the extraction of crude petroleum, the manufacture of tobacco products and of basic pharmaceutical goods (in terms of out-going links from the UK) ; the activities of head offices and holding companies, the sale of cars and manufacture of motor vehicles (in terms of in-going links to the UK). We showed in our experiment that although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalisation and thus can be expected to have a detrimental effect on the UK economy due to these foreign ownership links (see also S1 Text\ref{S1Text}). The impact of such unforeseen economic shocks could also be considered from a multi-level perspective, coupling our work with other approaches such as \cite{Dorretal2022} studying the firm-level impact of the Covid crisis.

@@

We also suggest that our simple agent-based model can be a basis for or a component of more complicated economic agent-based models. A co-evolution model with an evolution of city sizes in parallel to the firm network would be more realistic for an application on longer time scales, but would also be more difficult to parametrise and calibrate \cite{raimbault2018characterising}. A model with additional firm agents would lead to a multi-scale agent-based model, with also more flexibility for policy applications. The link with data could also be more elaborated: indeed the statistical analysis does not include spatial or neighborhood effects, and spatial non-stationarity may be accounted for using Geographically Weighted Regression models for example \cite{comber2021route}, to then be coupled with the ABM in an hybrid approach. Finally, an adaptation at the world scale using for example urban regions and data explored by \cite{rozenblat2021intra}, would be an extension with important practical and theoretical potentialities.

@@

\subsection*{S1 Text: UK's economic sectors dependent on international links.}
\label{S1Text}

@@

\begin{itemize}
    \item additional zero-inflated and hurdle statistical models;
    \item summary statistics for the socio-economic distance matrix;
    \item procedure for the synthetic model setup;
    \item Global Sensitivity Analysis
\end{itemize}

>>>>

\section*{Abstract}
The emergence of interconnected urban networks is a crucial feature of globalisation processes. Understanding the drivers behind the growth of such networks -- in particular urban firm networks --, is essential for the economic resilience of urban systems. We introduce in this paper a generative model for firm networks at the urban area level, which considers the following features: the economic size of the areas representing the source and target nodes, the industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distances between urban areas. An empirical network analysis on European urban firm ownership data confirms the relevance of each of these factors on new link creation. We then simulate network growth on synthetic systems of cities, unveiling stylised facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range. We calibrate the model on the European network, outperforming statistical models, and showing a strong role of path-dependency. The model is applied to stylised scenarios, opening up to the formulation of public policies recommendations to deal with exogenous shocks such as economic crises or potential lock-downs of countries.

@@

Networks constitute the new social morphology of our societies, since the intensification of globalisation of the economy in the late 20th century \cite{castells2000networksociety}. A world-city system has since then emerged \cite{taylor2001specification}, characterised by highly interconnected urban centres at the global level \cite{sassen1991}. These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity inducing spatial and non-spatial patterns \cite{martinus2018global}. Interactions of economic nature between transnational firms \cite{derudder2018central}, are together with international trade patterns among the most representative of the current economic trends \cite{taylor2001specification}. Understanding the drivers of growth and geographical properties of such an urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}. In addition, strengthening city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}. This can be achieved through strategic development of certain industrial sectors for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}. Exploring city/firm interactions \cite{storme2019introducing} and the position of cities within these networks \cite{gluckler2016relational} can also allow us to infer the consequences of unforeseen exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.

@@

At each time step, one directed link is randomly drawn. We define a generalised Cobb-Douglas utility function \cite{vilcu2011geometric} as

@@

The initial network is the same than in the synthetic setup, and the real links are taken as targets for the generated network. We constrain the total volume of flows by setting $w_0 = \sum_{ij} w_{ij}^{(obs)} / t_{f}$ when the number of time steps $t_f$ is given as a parameter and $w_{ij}^{(obs)}$ are observed flows.

@@

We base our study on empirical data from the \emph{AMADEUS} dataset produced by the Bureau Van Dijk (BVD) on transnational firms in Europe. \emph{AMADEUS} provides a precise location, turnover and activity codes for owners and owned companies in all sectors in Europe. The Bureau Van Dick gathers, homogenises and completes the information on these firms from national sources, publications of consultancies and audit firms such as KPMG, Deloitte, PricewaterhouseCoopers (PwC), Ernst \& Young (EY), Mazars, and those of other financial organisations such as Moody's, Fitch, Standard \& Poor's, Thomson Reuters or Duns and Bradstreet. A representativity study shows that \emph{ORBIS} --the Bureau Van Dijk's dataset for firms at the worldwide scale--, captures around 10 percent of existing foreign direct investments in 2011 in Poland \cite{Zdanowska2018}. \cite{bajgar2020coverage} find that the coverage of the \emph{ORBIS} database on OECD countries is a representative sample, but with a better performance on larger firms. As we aggregate company turnovers on urban areas in our analysis, we thus do not expect much representativity bias in the inter-urban network.

@@

A global sensitivity analysis first unveils an aggregated influence of parameters on model indicators. We apply Saltelli's Global Sensitivity method introduced by \cite{saltelli2008global}. This produces values for each parameter and each output of a first order sensitivity index, defined as $Var\left[\mathbb{E}_{\mathbf{X}_{\sim i}}\left(Y_j | X_i\right)\right] / Var\left[Y_j\right]$, for parameter $X_i$ and output $Y_j$, $\mathbf{X}_{\sim i}$ being all other parameters. This captures the total influence of $X_i$ all other things being equal. The total order index is given by $\mathbb{E}_{\mathbf{X}_{\sim i}} \left[Var(Y_j | \mathbf{X}_{\sim i})\right] / Var\left[Y_j\right]$ and captures non-homogeneity in behaviour and interaction between parameters. Indices were estimated with a Sobol sequence sampling of 20,000 parameter points (corresponding to 160,000 model runs) \cite{saltelli2010variance}. We give in S2 Text the full estimated results for sensitivity indices. To summarise, we find that indicators related to modularity (internationalisation and optimal modularity) are mostly influenced by geographical parameters. On the contrary, indicators of network structure are driven by origin and destination sizes. The influence of sectors and historical links has a secondary but not negligible role. Altogether, the analysis shows strong interaction effects between parameters, since for several indicators the total order index is one order of magnitude larger than the first order index. This means that the generative model captures some aspects of the complex behaviour of the system.

@@

We run a specific experiment to study the impact of the urban hierarchy on firms dynamics. Rank-size law in cities is a central stylised fact within urban theories, despite the discussions regarding the estimation of the power-law exponent, its actual values and its variation depending on the definition of cities considered \cite{cottineau2017metazipf,corral2020truncated}. I this particular context, understanding how this parameter can impact urban dynamics within artificial systems of cities -- in which it can be modified to extreme values non observable in real systems--, should help disentangling the variability due to model parameters from the variability due to the specific configuration of the urban system \cite{raimbault2019space}. We change the initial hierarchical exponent $\alpha$ from 0.1 to 2.0 with a step of 0.1, and also vary the interaction range $d_0$. The behaviour of the internationalisation and metropolisation indicators are shown in Fig.~\ref{fig:fig7}. As expected, the hierarchy of the urban system has an important impact on model outcomes. We find that more hierarchical urban systems tend to produce more international networks (recall that the internationalisation indicator corresponds to the modularity, the latter is minimised when the networks span more across different countries). This is consistent with the existence of globalised metropolis \cite{sassen1991global}. This effect is mitigated by reducing the interaction ranges. Regarding metropolisation, we observe the existence of maxima as a function of the initial hierarchy across different interaction ranges $d_0$. This means that a too high hierarchy is detrimental to the global metropolisation when considering all cities, what may be due to the fact that in such a case very large cities are capturing all flows, leaving nothing to intermediate and medium-sized cities. Interestingly, the optimal value $\alpha \simeq 1.1$ is not far from the observed values (which usually range between 0.9 and 1.1 but can take larger values such as 1.5 when extending the urban system \cite{Raimbault_2020}), implying that this optimality property is endogenous to the emergence of urban systems and linked to their intrinsic hierarchy.

%%%%%%%%%%%%%
\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{figures/Fig7.png}
    \end{center}
    \caption{\textbf{Impact of the urban hierarchy.} (Left) Internationalisation as a function of the initial synthetic hierarchical exponent $\alpha$, for different values of the interaction range $d_0$ (colour); (Right) Metropolisation as a function of $\alpha$, for different values of the interaction range $d_0$ (colour).\label{fig:fig7}}
\end{figure}
%%%%%%%%%%%%%

@@

We then apply the model to a real system of cities by calibrating it on the European ownership network. The model setup is done following the real setup described above. The number of time steps $t_f$ is left as an additional parameter, which in a sense determines the granularity of the cumulative network generation process. To avoid overfitting and including information from statistical models, we deactivate the socio-economic gravity interactions by setting $c_0 = 0$ (distances are obtained as fixed effects from a statistical model). Finally, $w_0$ can be set so that a given constraint is satisfied in a way similar to spatial interaction models \cite{wilson1975some}: in this case, we take an optimal value for the $\varepsilon_L$ objective given below as detailed in S2 Text.

@@

%%%%%%%%%%%%%%
\begin{figure}
	\begin{center}
    \includegraphics[width=0.75\linewidth]{figures/Fig8.png}
    \end{center}
	\caption{\textbf{Model calibration results.} Pareto front obtained for the bi-objective model calibration for $\varepsilon_L$ and $\varepsilon_M$ objectives. $\varepsilon_L$ can be compared to the MSE of statistical models fitted above, and the generative model outperform these for several parametrisations (lowest MSE 5.33 for statistical models). Point size gives the number of stochastic samples and colour the value of $\gamma_O$.\label{fig:fig8}}
\end{figure}
%%%%%%%%%%%%%%

@@

When considering optimal points with $\varepsilon_L < 5.0$, the adjusted parameters are: long range gravity interactions given by $d_0 = 7432 \pm 2350$; a stronger influence of the destination than of the origin as $\gamma_O = 1.41 \pm 0.11$ and $\gamma_D = 1.95 \pm 0.06$; a comparable influence of sectors as $\gamma_S = 1.20 \pm 0.14$; a fine time granularity as $t_f = 4927 \pm 119$; and a very strong influence of previous links as $\gamma_W = 7.72 \pm 5.15$. This last point furthermore shows in particular the importance of including path-dependency in this complex generative model. The interaction of this process with others implies an inversion of the relative importance of origin and destination (in opposition to the statistical models). The fact that calibrated interaction distance is large is consistent with long range economic interactions within Europe. Besides these stylised facts, these calibration results altogether demonstrate the feasibility of the application of the model to real data. As stated previously, the model is not predictive and the relatively high error values would not make it satisfactory in that manner, but are reasonable for its explanatory and application functions.

@@

%%%%%%%%%%%%%%
\begin{figure}
	\begin{center}
    \includegraphics[width=\linewidth]{figures/Fig9.png}
    \end{center}
	\caption{\textbf{Application of the calibrated model to an economic shock scenario.} All plots show the trajectory of indicators (internationalisation for the top row and average community size for the bottom row) in time, for different values of interaction decay rescaling $\kappa_G$ (colour) and socio-economic distance rescaling $\kappa_C$ (panels). The vertical dotted red line shows the moment the economic shock is introduced and $d_0$ (respectively $c_0$) is set to $\kappa_G \cdot d_0$ (resp. $\kappa_C \cdot c_0$).\label{fig:fig9}}
\end{figure}
%%%%%%%%%%%%%%

@@

The model calibration results provide model parametrisations reasonably approaching the observed network. This validates its potential application to stylised policy issues with practical applications in the future. For example the effect of exogenous shocks in the socio-economic structure can be tested as we did in the stylised application example above. Such shocks can for example be the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to further pandemic outbreaks on the single market. In this sense the UK economic sectors mostly dependent on foreign capital and generating the highest turnovers in 2018, that would be the most affected and worth examining in detail are: the extraction of crude petroleum, the manufacture of tobacco products and of basic pharmaceutical goods (in terms of out-going links from the UK) ; the activities of head offices and holding companies, the sale of cars and manufacture of motor vehicles (in terms of in-going links to the UK). We showed in our experiment that although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalisation and thus can be expected to have a detrimental effect on the UK economy due to these foreign ownership links (see also S1 Text). The impact of such unforeseen economic shocks could also be considered from a multi-level perspective, coupling our work with other approaches such as \cite{Dorretal2022} studying the firm-level impact of the Covid crisis.

@@

We also suggest that our simple agent-based model can be a basis for or a component of more complicated economic agent-based models. A co-evolution model with an evolution of city sizes in parallel to the firm network would be more realistic for an application on longer time scales, but would also be more difficult to parametrise and calibrate \cite{raimbault2018characterising}. A model with additional firm agents would lead to a multi-scale agent-based model, with also more flexibility for policy applications. The link with data could also be more elaborated: indeed the statistical analysis does not include spatial or neighbourhood effects, and spatial non-stationarity may be accounted for using Geographically Weighted Regression models for example \cite{comber2021route}, to then be coupled with the ABM in an hybrid approach. Finally, an adaptation at the world scale using for example urban regions and data explored by \cite{rozenblat2021intra}, would be an extension with important practical and theoretical potentialities.

@@

\subsection*{S1 Text: UK's economic sectors dependent on international links.\label{S1Text}}

@@

\begin{itemize}
    \item additional zero-inflated and hurdle statistical models;
    \item summary statistics for the socio-economic distance matrix;
    \item procedure for the synthetic model setup;
    \item Global Sensitivity Analysis;
    \item model calibration.
\end{itemize}
